\phantomsection\addcontentsline{toc}{chapter}{Acknowledgements} % Don't remove me!
\chapter*{\centering{\Large Acknowledgements}}

To my supervisors, Dr. Caren Han and Dr. Josiah Poon,

Thank you for the opportunity to work alongside you in the University of Sydney Natural Language Processing (NLP) research group. To Caren, I thank you for always making yourself available to answer my questions, reviewing my drafts and motivating me to keep developing new ideas. To Josiah, thank you for all your efforts in organising weekly research group meetings and providing wonderful opportunities to collaborate with other researchers despite the challenges of moving online this year.

\vspace{\baselineskip}

To my parents,

Thank you for your endless support over the past four years, for providing me with the opportunity to study full-time and for helping me to keep pushing forward though the many challenges I faced this year.

\vspace{\baselineskip}

To my girlfriend, Ayuka,

Thank you for standing by my side through the good and the bad, supporting me regardless of how my research was progressing throughout the year. Thank you for being the highlight of my day each and every day this year. I could not ask for a better friend and companion, and I am inexpressibly grateful to have you as a part of my life.

\vspace{\baselineskip}

To my sisters,

Thank you for always being generally cheerful and wonderful people to be around. It's been a pleasure to see you both approach the challenges of 2020 with such enthusiasm and success.

\vspace{\baselineskip}

To my friends, Sammy, Matt, Jessie and Ben,

Thank you for making the effort to stay in touch over what has been a difficult year. Though we haven't been able to meet in person for a while, I'm extremely grateful for our sporadic conversations over text and/or occasional video chats over a cup of coffee.
